---
title: "extra info"
format: html
editor: visual
embed-resources: true
---

necessary for regression

good to add a predictor variable. examples

```{r}
ggplot(features2,
       aes(x = length,
           fill = year)) +
  geom_density(alpha = .2,
               color = NA)
```

talk about limmitations - other variables

linear parallels slope - assumes length and rating are the same in both periods - so useless

alpha + Bleng + B2yr

scooped parallel - optimal movie length but still parallel - also assumes relationship same in both times

alpha + B1len + B2len2 + B3year

```{r}
m2 <- lm(rating ~ length + I(length^2) + year + 
           Action + Animation + Comedy + Drama + 
           Documentary + Romance,
         data = features2)
tidy(m2)

ggpredict(m2,
          terms = c("length", "year")) |> 
  plot()
```

Cross linear

alpha + B1len + B2year + B3(len)(yr)

```{r}
m3 <- lm(rating ~ length * year + 
           Action + Animation + Comedy + Drama + 
           Documentary + Romance,
         data = features2)
tidy(m3)

ggpredict(m3,
          terms = c("length", "year")) |> 
  plot()
```

should test and compare the 4 models
