---
title: "Homework 1"
author: "Devica Shah"
format: html
embed-resources: true
editor: visual
---

### Question 1

I installed the three packages. But I did it in the console as I was instructed.

### Question 2

```{r}
library(causact)
library(dplyr)
library(igraph)

df <- dplyr::as_data_frame(x = c(1,2,3))

glimpse(df)

n_distinct(x <- c(5,6,2,7,9,1))
```

In the initial call of "as_data_frame" which was "df \<- as_data_frame(x = c(1,2,3))", it produced an error because both the dplyr and igraph package have as_data_frame functions. Therefore the replaced line of code specifies a call to the as_data_frame function in the dplyr package to avoid errors.

When I ran the initial code it may have used the igraph function because of the order I loaded the packages. The igraph package was loaded after the dplyr package, so the igraph function may have overriden the dplyr one.

### Question 3

```{r}
?n_distinct()
```

I used the "?" to show the document of the help file for the n_distinct() function.

The n_distinct() function the number of distinct elements in a set of values or multiple sets of values.

### Question 4

```{r}
glimpse(baseballData)
```

I ran the glimpse function to look at the object baseballData and an organized array of information appeared.

The dataset has 12,145 rows and 5 columns.

The Home column has a "fct" variable which stands for factor.

The HomeScore column has an "int" variable which stands for integer.

### Question 5

```{r}
baseballData[1,]
```

After running the code, one row of the baseballData object appeared. It represents information about one baseball game.

```{r}
baseballData[,2:3] %>% head()
```

The two columns (column 2 and column 3) represent the acronym names of the home team and visitor team, respectively.

### Question 6

```{r}
name <-
  c(
    "Wayne Gretzky",
    "Gordie Howe",
    "Jaromir Jagr",
    "Brett Hull",
    "Marcel Dionne",
    "Phil Esposito" ,
    "Mike Gartner",
    "Alex Ovechkin",
    "Mark Messier" ,
    "Steve Yzerman"
  )

goals <- c(894, 801, 766, 741, 731, 717, 708, 700, 694, 692)

year_started <- c(1979, 1946, 1990, 1986, 1971, 1963, 1979, 2005, 1979, 1983)

hockeyPlayers <- tibble(
  name = name,
  goals = goals,
  year_started = year_started
)

glimpse(hockeyPlayers)
```

I created the dataframe titled hockeyPlayers using the tibble function. At first I was very confused but eventually after troubleshooting got it to work. I thought that each variable name needed to be in the format \<name\>, however realized that I need to take out the "\<\>".
