---
title: "Homework 03"
format: html
Author: Devica Shah
embed-resources: true
editor: visual
---

```{r}
library(tidyverse)
mario_kart <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/world_records.csv")
```

## Question 1

```{r}
three_laps <- mario_kart %>% 
  filter(type == "Three Lap")
```

I kept only the records for Three Laps as it said to do in the Homework.

```{r}
without_Rainbow_Road <- three_laps %>%
  filter(track != "Rainbow Road")
```

This is how I would create a dataset without records from the Rainbow Road Track. I used the "!=" logical operator to include all the tracks that do not equal "Rainbow Road" and saved them to a dataset titled "without_Rainbow_Road".

```{r}
only_Rainbow_Road <- three_laps %>%
  filter(track == "Rainbow Road")
```

I saved all the records on the track "Rainbow Road" to a dataset titled "only_Rainbow_Road".

## Question 2

```{r}
summarize_Rainbow_Road <- only_Rainbow_Road %>%
  summarize(mean = mean (time, na.rm = TRUE),
            std_dev = sd(time, na.rm = TRUE))
summarize_Rainbow_Road  
```

```{r}
summarize_three_laps <- three_laps %>%
  summarize(mean = mean (time, na.rm = TRUE),
            std_dev = sd(time, na.rm = TRUE))
summarize_three_laps
```

The rainbow road tracks has a higher average in time than the average time of all the tracks. Also, the rainbow road has a higher standard deviation than all the tracks, so the time data is more spread out for the rainbow road track.

## Question 3

```{r}
summarize_track <- three_laps %>%
  group_by(track) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
summarize_track
```

Toad's Turnpike has the most records established.

## Question 4

```{r}
summarize_driver <- three_laps %>%
  group_by(player,track) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
summarize_driver
```

The driver Penev had 26 records on Choco Mountain.

## Question 5

```{r}
average_track_time <- three_laps %>%
  group_by(track) %>%
  summarize(mean = mean(time, na.rum = TRUE)) %>%
  arrange(desc(mean))
average_track_time
```

Rainbow Road has the highest average time.

```{r}
three_laps_copy <- three_laps %>% 
  group_by(track) %>% 
  arrange(time) %>% 
  slice(1)  %>% 
  arrange(time) %>% 
  select(track, time)

three_laps_copy
```

## Question 6

```{r}
three_laps <- three_laps %>%
 mutate(longer_than_100 = if_else(record_duration > 100, 1L, 0L))

glimpse(three_laps)
```

```{r}
player_records <- three_laps %>%
  group_by(player) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
player_records
```

MR had the most long duration records.

## Question 7

```{r}
drivers <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/drivers.csv")
```

```{r}

drivers1 <- drivers %>%
   select(player,nation) %>%
   group_by(player) %>%
   slice_head(n = 1)

joined_drivers <-  left_join(three_laps, drivers1)
```

```{r}
ggplot(joined_drivers,
       aes(x = nation)) + geom_bar() 
```
